---
title: "Hw 6"
author: "Jordan Sutherland"
date: "3/23/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MASS)
library(VGAM)
```

## 1

The code below reads in the dataset.

```{r}
dataURL <- url("https://ajmolstad.github.io/docs/HW5_Q2.csv")
Smoking <- read.csv(dataURL, header=TRUE)
```

The code below fits a cumulative proportional odds model for the data.

```{r}
po.smoking <- vglm(breathing ~ factor(smoking) + age, family = cumulative(parallel = TRUE), data = Smoking, hde.NA = FALSE)
```

## 1a

The first line of code below fits a baseline-category logit model and is copied from the previous homework. The next lines of code calculate the AIC values of both models.

```{r}
bl.smoking <- vglm(breathing ~ factor(smoking) + age, family = multinomial, data = Smoking)
AIC(bl.smoking)
AIC(po.smoking)
```

AIC values can be used to compare whether this model fits the data better than the previous baseline-category logit model. The AIC values obtained from the models are similar. The original model has a slightly lower AIC which indicates a marginally better fit for the data. 

## 1b

The code below creates a non-proportional odds model and runs a likelihood ratio test for the proportional odds model.

```{r}
po.smoking2 <- vglm(breathing ~ factor(smoking) + age, family = cumulative(parallel = FALSE), data = Smoking, hde.NA = FALSE)
lrtest(po.smoking2, po.smoking)
```

The null hypothesis is that both beta11 and beta21 of the non-proportional odds model are equal to beta1 of the proportional odds model and similarly, beta12 and beta22 of the non-proportional odds model are equal to beta2 of the proportional odds model. The alternative hypothesis is one of the equalities is false. The test statistic is 5.5269. This follows a chi-squared distribution with 3 degrees of freedom under the null hypothesis.

With a p-value of 0.137, we fail to reject the null hypothesis. There is not enough evidence to conclude that either beta11 and beta21 of the non-proportional odds model are equal to beta1 of the proportional odds model or that beta12 and beta22 of the non-proportional odds model are equal to beta2 of the proportional odds model.

## 1c

The code below provides the estimates for the coefficients of the cumulative logit model with proportional odds.

```{r}
summary(po.smoking)
```

## 1d

```{r}
vcov(po.smoking)
```

## 2

The code below reads in the dataset.

```{r}
dataURL <- url("https://ajmolstad.github.io/docs/HW6_Q2.RDS")
Chemo <- readRDS(dataURL)
```

## 2a

```{r}
po.chemo <- vglm(response ~ treatment + gender, family = cumulative(parallel = TRUE), data = Chemo, hde.NA = FALSE)
summary(po.chemo)
```

There is evidence that the two treatment approaches have a different effect on disease progression since the treatmentsequential coefficient has a very small p-value. 

## 2b



## 2c

The code below creates a non-proportional odds model and runs a likelihood ratio test for the proportional odds model.

```{r}
po.chemo2 <- vglm(response ~ treatment + gender, family = cumulative(parallel = FALSE), data = Chemo, hde.NA = FALSE)
lrtest(po.chemo2, po.chemo)
```

The null hypothesis is that both beta11 and beta21 of the non-proportional odds model are equal to beta1 of the proportional odds model and similarly, beta12 and beta22 of the non-proportional odds model are equal to beta2 of the proportional odds model. The alternative hypothesis is one of the equalities is false. The test statistic is 3.1897 This follows a chi-squared distribution with 4 degrees of freedom under the null hypothesis.

With a very large p-value of 0.5266, we fail to reject the null hypothesis. There is not enough evidence to conclude that either beta11 and beta21 of the non-proportional odds model are equal to beta1 of the proportional odds model or that beta12 and beta22 of the non-proportional odds model are equal to beta2 of the proportional odds model.

## 3a

The response Y has 5 categories. Protestant is the most liberal while Other is the most conservative.

## 3b








