---
title: "STA4210 Project 3"
author: "Jordan Sutherland"
date: "4/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``1.`` The first line of code reads in the dataset and assigns names to the columns of Xi's for clarity on the variables. The next line of code provides a summary to check we are working with the correct dataset. Then, the R code divides the size variable by 10000 in order to gets the units correct and have it set up appropriately for regression analysis. The last R command displays the first three lines of the data.

```{r}
rental <- read.table("CH06PR18.txt",header=FALSE,
                     col.names=c("Y","age","expenses","vacrate","size"))
summary(rental)
rental$size <- rental$size/10000
rental[1:3,]
```

``2A.`` The first R command fits the linear model for the dataset in the order of X4, X1, X2, X3 (size, age, expenses, vacrate). The next line of code shows the ANOVA table for the linear model associated with the model that was coded above it.

```{r}
fit.4123 <- lm(Y ~ size + age + expenses + vacrate, data=rental)
anova(fit.4123)
```
``2B.`` The null hypothesis is that beta3=0 and the alternative hypothesis is that beta3 does not equal 0. The null distribution of the test statistic is F1,76. The test statistic value is Fobs= [SSR(X3|X1,X2,X4)/1]/[MSE(X1,X2,X3,X4)]=0.42/1.293=0.32. 

The first line of code gives the p-value of the test statistic calculated above and in the null distribution. The next line of code shows the minimum F statistic necessary to reject the null hypothesis at alpha level 0.01 with the null distribution.

```{r}
pf(0.32,1,76,lower.tail=FALSE)
qf(0.99,1,76)
```

With an F statistic of 0.32 < 6.98 and p-value 0.57 > 0.01 we fail to reject the null hypothesis at an alpha level of 0.01. There is not enough evidence to conclude that vacancy rate is a "useful" predictor and should be kept in the model after accounting for age, size, and expenses.

``2C.`` The null hypothesis is that beta2=beta3=0 and the alternative hypothesis is that at least one of beta2 and beta3 is not 0. The null distribution of the test statistic is F2,76. The test statistic value is Fobs= [SSR(X2,X3|X1,X4)/2]/[MSE(X1,X2,X3,X4)]
=[(SSR(X1,X2,X3,X4)-SSR(X1,X4))/2]/[SSR(X1,X2,X3,X4)/76]
=[((67.775+42.275+27.857+0.420)-(67.775+42.275))/2]/[98.231/76]
=14.1385/1.293=10.93.

The first line of code gives the p-value of the test statistic calculated above and in the null distribution. The next line of code shows the minimum F statistic necessary to reject the null hypothesis at alpha level 0.01 with the null distribution.

```{r}
pf(10.93,2,76,lower.tail=FALSE)
qf(0.99,2,76)
```

With an F statistic of 10.93 > 4.89 and p-value 6.73e-05 < 0.01, we reject the null hypothesis at an alpha level of 0.01. We have very strong evidence to conclude that at least one of the variables, expenses and vacancy rates, is "useful" to the model and should be kept in the model after accounting for age and size.

``3.`` The first R command calculates  mean of the age variable. The second line subtracts the mean from all age values in order to center age around the mean and makes it a new variable to the dataset: age.c. The next line fits the prediction equation for the dataset in order of X1, X2, X4, X11 (age.c, expenses, size, age.c^2). The next line of code shows the summary output for the linear model associated with the model that was coded above.

```{r}
agebar <- mean(rental$age)
rental$age.c <- rental$age - agebar
fit.q <- lm(Y ~ age.c + expenses +  size + I(age.c^2), data=rental)
summary(fit.q)
```

4. The test statistic I am using is Fobs=[SSR(X11|X1,X2,X4)/1]/[MSE(X1,X2,X4,X11)]
=7.115/1.204=5.91. The null distribution of the test statistic is F1,76.

The first line of code displays the ANOVA table for the fitted equation that contains the quadratic term. It can be used to calculate the F statistic since it shows the extra sums of squares that can be plugged into the formula. The second line of code gives the p-value of the test statistic calculated above and in the null distribution. The last line of code shows the minimum F statistic necessary to reject the null hypothesis at alpha level 0.01 with the null distribution.

```{r}
anova(fit.q)
pf(5.91,1,76,lower.tail=FALSE)
qf(0.99,1,76)
```

With an F statistic of 5.91 < 6.98 and p-value 0.0174 > 0.01, we fail to reject the null hypothesis at an alpha level of 0.01. There is not enough evidence to conclude that the quadratic term for age is a "useful" predictor and should be kept in the model after accounting for age, expenses, and size.

5. The t test statistic for the null hypothesis that beta11=0 is t=(0.014148-0)/0.005821=2.43. The t test statistic found corresonds with F=t^2 (5.91=2.43^2).

The first two lines of code set up the design matrix X with 1's in the left column and age.c, expenses, size, and age.c^2 in the next columns respectively. The next two lines solve for the inverse of X^(t)X. Then I set Y to be the vector of the Y values given in the dataset and calculated X^(t)Y. The last lines calculate betahat from the product of the inverse of X^(t)X and X^(t)Y.

```{r}
n <- nrow(rental)
X <- matrix(c(rep(1,n),rental$age.c,rental$expenses,rental$size,rental$age.c^2),nrow=n)
XtX <- t(X) %*% X; XtX
XtX.inv<-solve(XtX);round(XtX.inv,digits=2)
Y <- rental$Y
XtY <- t(X) %*% Y; XtY
betahat <- XtX.inv %*% XtY
betahat
```

The R code calculates the variance-covariance matrix of betahat from the sigma given in the summary output and the inverse matrix of X^(t)X calculated above.

```{r}
0.005821^2 * XtX.inv
```
